---
output: github_document
always_allow_html: true
---

<!-- This file is used to create README.md
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

```

```{r setup, include=FALSE}
library(tidyverse)
library(reportRmd)

```

# importExcel

importExcel is a department package to import data that has been checked with the internal UHN Biostatistics `dataChecker`. 

The package has one main function which will read in the excel data, add the labels and recode factors based on the information provided in the data dictionary. 

## Installation

You can install the latest version of importExcel from [GitHub](https://github.com/) with:


``` r
# install.packages("devtools")
devtools::install_github("biostatsPMH/importExcel")
```

## Documentation

[Online Documentation](https://biostatsPMH.github.io/importExcel/)

## Example 1: Tidy Data

This is a nice clean data set 


```{r example}
library(importExcel)
data_file <- system.file("extdata", "study_data.xlsx", package = "importExcel")

# This will read in the file, add variable labels and recode variables to 
# factors as specified in the dictionary sheet
file_import <- read_excel_with_dictionary(data_file = data_file,
                                          data_sheet = "data",dictionary_sheet = "dict")

# extract the recoded-data
coded_data <- file_import$coded_data

# file_import contains the following objects:
# # check warnings
# file_import$warnings
# 
# # check for text omitted from numeric data
# file_import$numeric_converstions
# 
# # check for attempted date conversions
# file_import$date_converstions
# 
# # Look at the dictionary used
# file_import$updated_dictionary

```

Describe the coded data:

```{r}
require(tidyverse)
require(reportRmd)

coded_data |> 
  select(!studyID) |> 
  rm_compactsum(xvars=everything())

```

## Example 2: Messier Data

This data has messy dates which are converted to yyyy-mm-dd format for consistency
```{r,warning=FALSE}

data_file <- system.file("extdata", "study_data_messy.xlsx", package = "importExcel")

# This will read in the file, add variable labels and recode variables to 
# factors as specified in the dictionary sheet
file_import <- read_excel_with_dictionary(data_file = data_file,
                                          data_sheet = "data",dictionary_sheet = "dict")

# extract the recoded-data
coded_data <- file_import$coded_data



```

Describe the coded data:

```{r}
coded_data |> 
  select(!studyID) |> 
  rm_compactsum(xvars=everything())

```

Look at the numeric values not imported:

```{r}
file_import$numeric_conversions
```

Look at the converted date values (not run): 

```{r, eval=F}
file_import$date_conversions
```


## Sample Data

The package installs two sample files: `study_data.xlsx` and `study_data_messy.xlsx`

The following code can be copied and run to view the first file:

```{r,eval=F}
tidy_data <- "study_data.xlsx"
messy_data <- "study_data_messy.xlsx"

file_path <- system.file("extdata", tidy_data, package = "importExcel")

# Check if the file exists
if (file.exists(file_path)) {
  # Open the file using the default system application (Excel)
  
  # For Windows
  if (Sys.info()['sysname'] == "Windows") {
    shell.exec(file_path)
  }
  
  # For macOS
  else if (Sys.info()['sysname'] == "Darwin") {
    system(paste("open", shQuote(file_path)))
  }
  
  # For Linux
  else if (Sys.info()['sysname'] == "Linux") {
    system(paste("xdg-open", shQuote(file_path)))
  }
  
  # If the OS is not recognized
  else {
    message("Unable to open file: Unrecognized operating system")
  }
} else {
  message("File 'study_data.xlsx' not found in the package's inst/extdata directory")
}

```
